# Components
## Input Parameters
### Essential Input parameters
For the input parameters, three options need to be defined usually, *id*, *type*, and *prefix*. The type can be *string*, *int*, *long*, *float*, *double*, and so on. More detail can be found at: <https://www.commonwl.org/v1.0/CommandLineTool.html#CWLType>.

Here is an example from [CWL user guide](http://www.commonwl.org/user_guide/03-input/). Here we defined an `echo` with different type of input parameters by `InputParam`. The `stdout` option can be used to caputre the standard output stream to a file.

```{r}
e1 <- InputParam(id = "flag", type = "boolean", prefix = "-f")
e2 <- InputParam(id = "string", type = "string", prefix = "-s")
e3 <- InputParam(id = "int", type = "int", prefix = "-i")
e4 <- InputParam(id = "file", type = "File", prefix = "--file=", separate = FALSE)
echoA <- cwlParam(baseCommand = "echo",
                  inputs = InputParamList(e1, e2, e3, e4),
                  stdout = "output.txt")
```

Then we give it a try by setting values for the inputs.
```{r}
echoA$flag <- TRUE
echoA$string <- "Hello"
echoA$int <- 1

tmpfile <- tempfile()
write("World", tmpfile)
echoA$file <- tmpfile

r2 <- runCWL(echoA, outdir = tempdir())
r2$command
```

### Array Inputs
A similar example to CWL user guide. We can define three different type of array as inputs. 
```{r}
a1 <- InputParam(id = "A", type = "string[]", prefix = "-A")
a2 <- InputParam(id = "B",
                 type = InputArrayParam(items = "string",
                                        prefix="-B=", separate = FALSE))
a3 <- InputParam(id = "C", type = "string[]", prefix = "-C=",
                 itemSeparator = ",", separate = FALSE)
echoB <- cwlParam(baseCommand = "echo",
                 inputs = InputParamList(a1, a2, a3))
```

Then set values for the three inputs.
```{r}
echoB$A <- letters[1:3]
echoB$B <- letters[4:6]
echoB$C <- letters[7:9]
echoB
```

Now we can check whether the command behaves as we expected.
```{r}
r3 <- runCWL(echoB, outdir = tempdir())
r3$command
```

## Output Parameters
### Capturing Output
The outputs, similar to the inputs, is a list of output parameters. Three options *id*, *type* and *glob* can be defined. The glob option is used to define a pattern to find files relative to the output directory.

Here is an example to unzip a compressed `gz` file. First, we generate a compressed R script file.
```{r}
zzfil <- file.path(tempdir(), "sample.R.gz")
zz <- gzfile(zzfil, "w")
cat("sample(1:10, 5)", file = zz, sep = "\n")
close(zz)
```

We define a `cwlParam` object to use "gzip" to uncompress a input file.
```{r}
ofile <- "sample.R"
z1 <- InputParam(id = "uncomp", type = "boolean", prefix = "-d")
z2 <- InputParam(id = "out", type = "boolean", prefix = "-c")
z3 <- InputParam(id = "zfile", type = "File")
o1 <- OutputParam(id = "rfile", type = "File", glob = ofile)
gz <- cwlParam(baseCommand = "gzip",
               inputs = InputParamList(z1, z2, z3),
               outputs = OutputParamList(o1),
               stdout = ofile)
```

Now the `gz` object can be used to uncompress the previous generated compressed file.
```{r}
gz$uncomp <- TRUE
gz$out <- TRUE
gz$zfile <- zzfil
r4 <- runCWL(gz, outdir = tempdir())
r4$output
```

Or we can use `arguments` to set some default parameters.
```{r}
z1 <- InputParam(id = "zfile", type = "File")
o1 <- OutputParam(id = "rfile", type = "File", glob = ofile)
Gz <- cwlParam(baseCommand = "gzip",
               arguments = list("-d", "-c"),
               inputs = InputParamList(z1),
               outputs = OutputParamList(o1),
               stdout = ofile)
Gz
Gz$zfile <- zzfil
r4a <- runCWL(Gz, outdir = tempdir())
```

To make it for general usage, we can define a pattern with javascript to glob the output, which require `node` to be installed in your system PATH.
```{r}
pfile <- "$(inputs.zfile.path.split('/').slice(-1)[0].split('.').slice(0,-1).join('.'))"
```

Or we can use the CWL built in file property, `nameroot`, directly.
```{r}
pfile <- "$(inputs.zfile.nameroot)"
o2 <- OutputParam(id = "rfile", type = "File", glob = pfile)
req1 <- list(class = "InlineJavascriptRequirement")
GZ <- cwlParam(baseCommand = c("gzip", "-d", "-c"),
               requirements = list(), ## assign list(req1) if node installed.
               inputs = InputParamList(z1),
               outputs = OutputParamList(o2),
               stdout = pfile)
GZ$zfile <- zzfil
r4b <- runCWL(GZ, outdir = tempdir())
```

### Array Outputs

We can also capture multiple output files with `glob` pattern.
```{r}
a <- InputParam(id = "a", type = InputArrayParam(items = "string"))
b <- OutputParam(id = "b", type = OutputArrayParam(items = "File"), glob = "*.txt")
touch <- cwlParam(baseCommand = "touch", inputs = InputParamList(a), outputs = OutputParamList(b))
touch$a <- c("a.txt", "b.gz", "c.txt")
r5 <- runCWL(touch, outdir = tempdir())
r5$output
```
