# Writing Pipeline
We can connect multiple tools into a pipeline. Here is an example to uncompress an R script and execute it with `Rscript`.

We first define a simple `Rscript` tool without using docker:
```{r}
d1 <- InputParam(id = "rfile", type = "File")
Rs <- cwlParam(baseCommand = "/usr/bin/Rscript",
               inputs = InputParamList(d1))
Rs
```

Test run:
```{r}
Rs$rfile <- r4$output
tres <- runCWL(Rs, outdir = tempdir())
readLines(tres$output)
```

The pipeline includes two steps, decompressed by `GZ` and compiled by `Rs`. The input file is a compressed file and the output would be the output `Rout` from `Rs`.

First we need to define the direct inputs and outputs from `GZ` and `Rs`:
```{r}
i1 <- InputParam(id = "cwl_zfile", type = "File")
o1 <- OutputParam(id = "cwl_cout", type = "File", outputSource = "Compile/output")
```
For the input "cwl_zifle", it refers to the `GZ` input `zfile`.  The output "cwl_cout" will be the outcome of `Rs` output `Rout`.

The `cwlStepParam` is used to define `inputs` and `outputs` from previous step. Then it connects with the two steps with `Step` function. The `run` option refer to the corresponding `cwlParam` object and the `In` option in steps should be linked to the input parameters defined by `cwlStepParam`. In the end, we use `+` to connect all steps.
```{r}
cwl <- cwlStepParam(inputs = InputParamList(i1),
                    outputs = OutputParamList(o1))
s1 <- Step(id = "Uncomp", run = GZ,
           In = list(zfile = "cwl_zfile"))
s2 <- Step(id = "Compile", run = Rs,
           In = list(rfile = "Uncomp/rfile"))
cwl <- cwl + s1 + s2
cwl
```

Let's run the pipeline.
```{r}
cwl$cwl_zfile <- zzfil
r7 <- runCWL(cwl, outdir = tempdir())
readLines(r7$output)
```

## Scattering pipeline
The scattering feature can specifies the associated workflow step or subworkflow to execute separately over a list of input elements. To use this feature, `ScatterFeatureRequirement` must be specified in the workflow requirements. Different `scatter` methods can be used in the associated step to decompose the input into a discrete set of jobs. More details can be found at: https://www.commonwl.org/v1.0/Workflow.html#WorkflowStep.

Here is an example to execute multiple R scripts. First, we need to set the input and output types to be array of "File", and add the requirments. In the "Compile" step, the scattering input is required to be set with the `scatter` option.
```{r}
i2 <- InputParam(id = "cwl_rfiles", type = "File[]")
o2 <- OutputParam(id = "cwl_couts", type = "File[]", outputSource = "Compile/output")
req1 <- list(class = "ScatterFeatureRequirement")

cwl2 <- cwlStepParam(requirements = list(req1),
                     inputs = InputParamList(i2),
                     outputs = OutputParamList(o2))
s1 <- Step(id = "Compile", run = Rs,
           In = list(rfile = "cwl_rfiles"),
           scatter = "rfile")
cwl2 <- cwl2 + s1
cwl2
```

Multiple R scripts can be assigned to the workflow inputs and executed.
```{r}
cwl2$cwl_rfiles <- c(r4b$output, r4b$output)
r8 <- runCWL(cwl2, outdir = tempdir())
r8$output
```

## Pipeline plot
The function `plotCWL` can be used to visualize the relationship of inputs, outputs and the analysis for a tool or pipeline.

```{r}
plotCWL(cwl)
```
